---
title: "Practicum_Two"
output:
  pdf_document: default
  html_notebook: default
---
# Create database schema
## Connect to database
```{r}
library(RSQLite)

dbFile = "./Practicum2.sqlite"

dbcon <- dbConnect(RSQLite::SQLite(), dbFile)
```

## Create table Author
```{sql connection=dbcon}
DROP TABLE IF EXISTS Author;
```

```{sql connection=dbcon}
CREATE TABLE Author (
  authorId INTEGER PRIMARY KEY,
  lastName TEXT,
  foreName TEXT,
  initials TEXT
);
```

## Create table Journal
```{sql connection=dbcon}
DROP TABLE IF EXISTS Journal;
```

```{sql connection=dbcon}
CREATE TABLE Journal (
  ISSN TEXT PRIMARY KEY,
  title TEXT NOT NULL,
  isoabbreviation TEXT NOT NULL
);
```

## Create table JournalIssue
```{sql connection=dbcon}
DROP TABLE IF EXISTS JournalIssue;
```

```{sql connection=dbcon}
CREATE TABLE JournalIssue (
  jId INTEGER PRIMARY KEY,
  ISSN INTEGER REFERENCES Journal(ISSN) ON DELETE CASCADE,
  volume INTEGER NOT NULL,
  issue INTEGER NOT NULL,
  pubYear TEXT NOT NULL,
  pubMonth TEXT NOT NULL
);
```

## Create table AuthorShip
```{sql connection=dbcon}
DROP TABLE IF EXISTS AuthorShip;
```

```{sql connection=dbcon}
CREATE TABLE AuthorShip (
  articleId INTEGER,
  authorId INTEGER,
  PRIMARY KEY (articleId, authorId)
)
```

## Create table Article
```{sql connection=dbcon}
DROP TABLE IF EXISTS Article;
```

```{sql connection=dbcon}
CREATE TABLE Article (
  articleId INTEGER PRIMARY KEY,
  jId INTEGER REFERENCES JournalIssue(jId) ON DELETE CASCADE,
  pubMode TEXT NOT NULL,
  title TEXT NOT NULL,
  abstract TEXT NOT NULL
)
```

## Create table History
```{sql connection=dbcon}
DROP TABLE IF EXISTS History;
```

```{sql connection=dbcon}
CREATE TABLE History (
  hId INTEGER PRIMARY KEY,
  articleId INTEGER REFERENCES Article(articleId) ON DELETE CASCADE,
  pubStatus TEXT NOT NULL,
  year INTEGER NOT NULL,
  month INTEGER NOT NULL,
  day INTEGER NOT NULL
)
```

# Loading Libraries

```{r warning=FALSE}
library(XML)
library(xslt)
```

# Load XML data to database
## Read XML file
```{r}
xmlPath <- "./pubmed_sample.xml"
xmlDoc = read_xml(xmlPath)
```

## Load Journal table
```{r}
# Extract Journal Data
xslPath = "./journal.xsl"
xslDoc = read_xml(xslPath,package="xslt")
journal_raw = xml_xslt(xmlDoc,xslDoc)
journals = unique(xmlToDataFrame(XML::xmlParse(journal_raw)))
dbWriteTable(dbcon,name = "Journal", value = journals, append = TRUE)
```

## Load Author table
```{r}
# Extract Author Data
xslPath = "./author.xsl"
xslDoc = read_xml(xslPath,package="xslt")
author_raw = xml_xslt(xmlDoc,xslDoc)
authors = unique(xmlToDataFrame(XML::xmlParse(author_raw)))
authors = cbind(authorId=rownames(authors),authors)
dbWriteTable(dbcon,name = "Author", value = authors, append = TRUE)
```

```{sql connection=dbcon}
SELECT * FROM Author;
```
## Load JournalIssue table
```{r}
# Extract JournalIssue Data
xslPath = "./journalIssue.xsl"
xslDoc = read_xml(xslPath,package="xslt")
journalIssue_raw = xml_xslt(xmlDoc,xslDoc)
journalIssues = unique(xmlToDataFrame(XML::xmlParse(journalIssue_raw)))
journalIssues = cbind(jId=rownames(journalIssues),journalIssues)
medDate <- journalIssues$medlineDate
for (i in 1 : nrow(journalIssues)) {
  if (medDate[i] != '') {
  pubDate <- strsplit(medDate[i],split = ' ')
  journalIssues$pubYear[i] = pubDate[[1]][1]
  journalIssues$pubMonth[i] = pubDate[[1]][2]
  }
}
journalIssues <- subset(journalIssues, select = -c(7))
dbWriteTable(dbcon,name = "JournalIssue", value = journalIssues, append = TRUE)
```

```{sql connection=dbcon}
SELECT * FROM JournalIssue;
```

## Load Article table
```{r}
# Extract Article Data
xslPath = "./Article.xsl"
xslDoc = read_xml(xslPath,package="xslt")
article_raw = xml_xslt(xmlDoc,xslDoc)
article = unique(xmlToDataFrame(XML::xmlParse(article_raw)))
rownames(article) <- NULL
jIds = c()
for(i in 1:nrow(article)){
  temp = article[i,]
  jIds = append(jIds,journalIssues[which(temp$ISSN==journalIssues$ISSN
                       & temp$volume==journalIssues$volume
                       & temp$issue == journalIssues$issue),]$jId)
}
article = cbind(articleId=rownames(article),jId=jIds,article)
article = subset(article, select = 1:5)
dbWriteTable(dbcon,name = "Article", value = article, append = TRUE)
```

```{sql connection=dbcon}
SELECT * FROM Article;
```


## Load AuthorShip table
```{r}
# Extract authorship Data
xslPath = "./AuthorShip.xsl"
xslDoc = read_xml(xslPath,package="xslt")
authorship_raw = xml_xslt(xmlDoc,xslDoc)
authorship = unique(xmlToDataFrame(XML::xmlParse(authorship_raw)))
rownames(authorship) <- NULL
authorIds = c()
articleIds = c()
for(i in 1:nrow(authorship)){
  temp = authorship[i,]
  articleIds = append(articleIds, article[which(temp$title==article$title),]$articleId)
  authorIds = append(authorIds,authors[which(temp$LastName==authors$lastName
                       & temp$ForeName==authors$foreName),]$authorId)
}
authorship = cbind(articleId=articleIds,authorId=authorIds,authorship)
authorship = subset(authorship, select = 1:2)
dbWriteTable(dbcon,name = "Authorship", value = authorship, append = TRUE)
```

```{sql connection=dbcon}
SELECT * FROM AuthorShip;
```


## Load History table
```{r}
# Extract History Data
xslPath = "./history.xsl"
xslDoc = read_xml(xslPath,package="xslt")
history_raw = xml_xslt(xmlDoc,xslDoc)
historys = unique(xmlToDataFrame(XML::xmlParse(history_raw)))
rownames(historys) <- NULL
articleIds = c()
for(i in 1:nrow(historys)){
  temp = historys[i,]
  articleIds = append(articleIds,article[which(temp$title==article$title),]$articleId)
}
historys = cbind(hId=rownames(historys),articleId=articleIds,historys)
historys = subset(historys, select = 1:6)
dbWriteTable(dbcon,name = "History", value = historys, append = TRUE)
```

```{sql connection=dbcon}
SELECT * FROM History;
```

**q2**

```{r}
sqlCmd <- "SELECT ji.jId AS jId, j.ISSN AS ISSN, j.title AS title, 
              ji.volume AS volume, ji.issue AS issue, 
              ji.pubYear As pubYear, ji.pubMonth As pubMonth
            FROM Journal AS j
            JOIN JournalIssue AS ji ON j.ISSN = ji.ISSN"
journal = dbGetQuery(dbcon, sqlCmd)
```

```{r}
sqlCmd <- "SELECT authorID, lastName, foreName
            FROM Author"
author = dbGetQuery(dbcon, sqlCmd)
```

```{r}
sqlCmd <- "SELECT hId, pubStatus, year, month, day
            FROM History"
history = dbGetQuery(dbcon, sqlCmd)
```

```{r}
sqlCmd <- "SELECT JournalIssue.jId AS jId, AuthorShip.authorId AS authorId, 
            History.hId AS hId, JournalIssue.ISSN AS ISSN, 
            Article.pubMode AS pubMode, Article.title AS title
          FROM Article
          JOIN AuthorShip ON Article.articleId = AuthorShip.articleId
          JOIN JournalIssue ON Article.jId = JournalIssue.jId
          JOIN History ON Article.articleId = History.articleId"
temp = dbGetQuery(dbcon,sqlCmd)
```


```{r}
library(RSQLite)
library(ggplot2)

dbFile = "./Practicum2_Part2.sqlite"

dbcon1 <- dbConnect(RSQLite::SQLite(), dbFile)
```

## Create table Journals
```{sql connection=dbcon1}
DROP TABLE IF EXISTS Journals;
```

```{sql connection=dbcon1}
CREATE TABLE Journals (
  jId INTEGER PRIMARY KEY,
  ISSN TEXT NOT NULL,
  volume INTEGER NOT NULL,
  issue INTEGER NOT NULL,
  pubYear TEXT NOT NULL,
  pubMonth TEXT NOT NULL,
  title TEXT NOT NULL
);
```

## Create table Author
```{sql connection=dbcon1}
DROP TABLE IF EXISTS Authors;
```

```{sql connection=dbcon1}
CREATE TABLE Authors (
  authorId INTEGER PRIMARY KEY,
  lastName TEXT,
  foreName TEXT
);
```

## Create table History
```{sql connection=dbcon1}
DROP TABLE IF EXISTS Historys;
```

```{sql connection=dbcon1}
CREATE TABLE Historys (
  hId INTEGER PRIMARY KEY,
  pubStatus TEXT NOT NULL,
  year INTEGER NOT NULL,
  month INTEGER NOT NULL,
  day INTEGER NOT NULL
)
```

## Create table ArticleFact
```{sql connection=dbcon1}
DROP TABLE IF EXISTS ArticleFact;
```

```{sql connection=dbcon1}
CREATE TABLE ArticleFact (
  jId TEXT NOT NULL,
  authorId INTEGER NOT NULL,
  hId INTEGER NOT NULL,
  pubMode TEXT,
  title TEXT,
  PRIMARY KEY (jId, authorId, hId),
  FOREIGN KEY (jId) REFERENCES Journals (jId),
  FOREIGN KEY (authorId) REFERENCES Authors (authorId),
  FOREIGN KEY (hId) REFERENCES Historys (hId)
);
```



```{r}
dbWriteTable(dbcon1, "Journals", journal, overwrite = T, row.names = F)
dbWriteTable(dbcon1, "Authors", author, overwrite = T, row.names = F)
dbWriteTable(dbcon1, "Historys", history, overwrite = T, row.names = F)
```

```{sql connection=dbcon1}
SELECT * FROM Journals
```

```{sql connection=dbcon1}
SELECT * FROM Authors
```

```{sql connection=dbcon1}
SELECT * FROM Historys
```

```{r}
patternData <- dbGetQuery(dbcon, "select month, count(1) as hid from History where pubstatus = 'pubmed' group by month order by month asc")
month <- patternData[,1]
publications <- patternData[,2]

ggplot(patternData, aes(x=month, y=publications)) + geom_line() + xlim(0,12) + ylim(0,6) + labs(title="Monthly Pattern for Publication") + theme(plot.title = element_text(hjust = 1)) + scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10,11,12))

```


```{r}
dbDisconnect(dbcon)
dbDisconnect(dbcon1)
```


